A Bayesian ranking statistic to find \\high-mass black holes in LIGO data

### 2021-01-28 main.tex
add plotting scripts
Updates from Overleaf

Until recently, there was no definitive evidence for intermediate-mass black holes, the black holes whose masses lie between that of stellar-mass and supermassive black hole systems ($10^2-10^6\ \msun$). This changed with the detection of GW190521, a unique gravitational wave event that led to the formation of a black hole with a mass  $142\ \msun$, the first confirmed direct discovery of an intermediate-mass black hole~\cite{gw190521}. Although this is the first gravitational wave event leading to discovering a black hole with a mass in the intermediate range, ground-based gravitational wave detectors are sensitive to gravitational waves from even more massive systems to systems with a total mass of $400\ \msun$. \av{double check the following -- is this the most recent rate quote?} Additionally, gravitational waves from systems with masses greater than $100\ \msun$ systems should occur at a rate of $0-10\text{yr}^{-1}$~\cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions,ligo_imbh_search}. Hence, LIGO might have gravitational wave events from more intermediate-mass systems hidden in its data, potentially even in O2 data.

@@

\begin{figure}[!ht]

{\centering \includegraphics[width=0.75\linewidth]{images/template_bank_masses}

}
% time for chunk 14
% 1176955218, Apr 23, 2017, 4:00 UTC
% 1178294418, May 08, 2017, 16:00 UTC
\caption[High-mass BCR search space.]{The template bank used by \pycbc to search a section of O2 data from $\text{April 23 - May 8, 2017}$. Our search is constrained to the high-mass parameter space enclosed by the dashed line. \av{Update plot }}\label{fig:templateBank}
\end{figure}

@@

\begin{figure}[!ht]
{\centering \includegraphics[width=0.75\linewidth]{images/bcr_cdf} }
\caption[BCR distribution example]{Histograms represent the survival function (1-CDF) from our selection of $\sim$ 3,000 background triggers (grey) and 648 simulated signals (blue) triggers obtained from \pycbc's search of data from $\text{April 23 - May 8, 2017}$. \av{change BCR to\bcr in axes labels}Vertical lines mark the ln BCRs of two glitches (orange and yellow), IAS's GW170121 (pink), and GWTC-1's GW170104 (dark blue) \av{this plot also doesn't add much, is it needed?}.}\label{fig:bcrCdf}
\end{figure}

@@

O2 lasted several months over which the detector's sensitivity varied. Hence, a part of our analysis entailed tuning the prior odds for obtaining a signal and a glitch, $\alpha$ and $\beta$, as described in Section~\ref{sec:method}. Table~\ref{tab:priorodds} presents the signal and glitch prior odds utilised for each time-frame of O2 data.
\input{Data/alpha_beta_table}

Tuning the prior odds can dramatically affect the \pastrobcr. For example, consider Table~\ref{tab:tuningresults}, which reports tuned \pastrobcr and un-tuned \untunedpastrobcr (where $\alpha=1$ and $\beta=1$) for various high-mass events and candidates. By tuning the prior odds, the \pastrobcr for some IAS events (for example, GW170403 and GW170817A) can change by more than 0.5, resulting in the promotion/demotion of a candidate's significance.
\input{Data/tuning_results_table}

>>>>

Until recently, there was no definitive evidence for intermediate-mass black holes, the black holes whose masses lie between that of stellar-mass and supermassive black hole systems ($10^2-10^6\ \msun$). This changed with the detection of GW190521, a unique gravitational wave event that led to the formation of a black hole with a mass  $142\ \msun$, the first confirmed direct discovery of an intermediate-mass black hole~\cite{gw190521}. Although this is the first gravitational wave event leading to discovering a black hole with a mass in the intermediate range, ground-based gravitational wave detectors are sensitive to gravitational waves from even more massive systems to systems with a total mass of $400\ \msun$~\cite{GWTC1}. Additionally, gravitational waves from systems with masses greater than $100\ \msun$ systems should occur at a rate of $0-10\text{yr}^{-1}$~\cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions,ligo_imbh_search}.  \av{this is surely outdated... get more recent rate quote} Hence, LIGO might have gravitational wave events from more intermediate-mass systems hidden in its data, potentially even in O2 data.

@@

\begin{figure*}[!ht]

{\centering \includegraphics[width=0.75\linewidth]{images/template_bank.png}

}
% time for chunk 14
% 1176955218, Apr 23, 2017, 4:00 UTC
% 1178294418, May 08, 2017, 16:00 UTC
\caption[High-mass BCR search space.]{The template bank (pink) used by \pycbc to search a section of O2 data from $\text{April 23 - May 8, 2017}$. Our search is constrained to the high-mass parameter space enclosed by the dashed line. The candidate, background and simulated triggers detected in this region of the parameter space during this period are plotted in orange, grey and blue respectively.  }\label{fig:templateBank}
\end{figure*}

@@

\begin{figure}[!ht]
{\centering \includegraphics[width=0.75\linewidth]{images/bcr_cdf_smaller_legend.png} }
\caption[BCR distribution example]{Histograms represent the survival function (1-CDF) from our high-mass selection of background triggers (grey) and simulated signals (blue) triggers obtained from \pycbc's search of data from $\text{August 13 - 21, 2017}$. Vertical lines mark the $\text{ln}\ \bcr$ of IAS's GW170817 and GWTC-1's GW170814.}\label{fig:bcrCdf}
\end{figure}

@@

O2 lasted several months over which the detector's sensitivity varied. Hence, a part of our analysis entailed tuning the prior odds for obtaining a signal and a glitch, $\alpha$ and $\beta$, as described in Section~\ref{sec:method}. Table~\ref{tab:priorodds} presents the signal and glitch prior odds utilised for each time-frame of O2 data. \av{note: havent tuned all chunks yet -- and tbh the 'tuning' could be better, atm brute force searching a 2d log-spce grid}
\input{Data/alpha_beta_table}

Tuning the prior odds can dramatically affect the \pastrobcr. For example, consider Table~\ref{tab:tuningresults}, which reports tuned \pastrobcr and un-tuned \untunedpastrobcr (where $\alpha=1$ and $\beta=1$) for various high-mass events and candidates. By tuning the prior odds, the \pastrobcr for some IAS events (for example, GW170403 and GW170817A) can change by more than 0.5, resulting in the promotion/demotion of a candidate's significance.

\input{Data/tuning_results_table}
